---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?

A1. We specified the amounts and after running the code, we found out that how many contributions made using each category. For instance, there are 5674 contributions made between $10-25. 

```{r}
"wv_winred" <- wv_winred_contribs  %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount < 26 ~ "Between $10 and $25",
        amount >= 26 & amount < 51 ~ "Between $26 and $50",
        amount >= 51 & amount < 100 ~ "Between $51 and $100",
        amount >= 101 & amount < 500 ~ "Between $101 and $500",
        amount >= 500 ~ "More than $500"
      ) )%>% 

    group_by(amount_category, date) %>%
    summarize (count_contribs = n())
```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  

A2. There are 737 donors who contributed less than $10. Dr. Rev John Almond contributed 293 times which is more than others in the list. Franco Colonna is the second top who contributed 183 times.  

```{r}
wv_winred_10 <- wv_winred <- wv_winred_contribs  %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount < 26 ~ "Between $10 and $25",
        amount >= 26 & amount < 51 ~ "Between $26 and $50",
        amount >= 51 & amount < 100 ~ "Between $51 and $100",
        amount >= 101 & amount < 500 ~ "Between $101 and $500",
        amount >= 500 ~ "More than $500"
      ) )%>% 
  filter(amount_category == "Less than $10")  %>% 

  group_by(last_name, first_name, address_one) %>%
    summarize (count_contribs = n()) %>% 
    arrange(desc(count_contribs))
```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. To be honest, I first got confused about coding and how even after I used filter, I still got numbers above $10. Later however, when I realized the number is not the amount and it is the number of times each person contributed, I got surprised. What made each of these donors to contribute more than once? How and why this happened? 

Q4. Which month saw the most contributions?
A4. If you meant, contributions related to "less than $10", then the answer is April and the num of contribution is 34. 

how to figure the answer for the question you asked about each price category that you give in the first question. 

```{r}
wv_winred_10 <- wv_winred <- wv_winred_contribs  %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount < 26 ~ "Between $10 and $25",
        amount >= 26 & amount < 51 ~ "Between $26 and $50",
        amount >= 51 & amount < 100 ~ "Between $51 and $100",
        amount >= 101 & amount < 500 ~ "Between $101 and $500",
        amount >= 500 ~ "More than $500"
      ) )%>% 
  filter(amount_category == "Less than $10")  %>% 

  group_by(last_name, first_name, address_one, date) %>%
    summarize (count_contribs = n()) %>% 
    arrange(desc(count_contribs))
```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.

A5. In total, there are six and all are from WV. I found all except for one who I have doubt.

```{r}

wv_winred_10 <- wv_winred <- wv_winred_contribs  %>%

  filter(amount == 2900)  %>% 

  group_by(last_name, first_name, date, amount) %>%
    summarize (count_contribs = n()) %>%
    arrange(desc(count_contribs))
```
