---
title: "lab_05"
author: "Derek Willis"
date: "8/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
library(refinr)

```

For this lab, we want to investigate spending by Maryland state and local candidates during the current election. For example, we could ask how much money have candidates and committees spent on advertising on Facebook? Which vendors received the most money for media expenses? We have the data, but it's messy - names of payees are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis.

## Load Data

You'll need to load one data set, a CSV file of Maryland campaign expenditures located in the data folder called "md_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to clean up the column names and make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table and make any fixes/changes you need to

# Load data
md_expenses <- read_csv("data/md_expenses.csv")

# Display it
md_expenses
```

## Answer questions

**Q1.** You've been assigned to write about Maryland political committees' use of Facebook advertising. There's a category column in the data that has values for different kinds of spending. One of them is "Media". Write code to isolate Media expenses where the name of the payee is not NA. Then, using `write_csv`, make a CSV file in your data folder. 

Using OpenRefine, create a project using that CSV file, create a copy of the `payee_name` column called `payee_clean` and then standardize it, focusing on the payees that appear most often. Pay particular attention to variations of Facebook, bearing in mind that it changed its name last year: https://about.fb.com/news/2021/10/facebook-company-is-now-meta/), but also clean up variations of the most common payees.

Then, when you are done standardizing `payee_clean`, export it as a new CSV file to your data folder, *giving it a different name* and read it back into RStudio as a new dataframe.

Then, using that `payee_clean` column, write code to find all of the spending where Facebook is the payee and calculate the total amount of spending for each committee and present the results showing the committee that spent the most first.

**A1.**  Jain  Ashwani Friends of (62357.38), Franchot Anderson-Walker for Maryland Slate (57928.25), Franchot  Peter Friends Of (50953.81), Boniface  Billy Friends Of (17705.47), Baron  Jon (Jonathan) for Maryland (15794.1)

```{r}
Expense <- md_expenses %>%
   filter(`Expense Category` == "Media")
     head(Expense)  
    
write_csv(Expense, "Expenses.csv")
  
```

```{r}
payee_fb <- read_csv("data/payee_facebook.csv") 

```
```{r}
Payee_FB <- payee_fb %>% 
 filter(payee_clean == "Facebook") %>% 
group_by(`Committee Name`) %>% 
summarise(total_amount=sum(`Amount($)`)) %>% 
  arrange(desc(total_amount))
```

**Q2.** Let's do some research on the top committees from question 1. Look up the top 10 spenders: if they are candidate committees, what offices were they running for? If they were not candidates, what issues do they seem to care about? Write some sentences answering those questions, then go the Facebook ad library (https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data[direction]=desc&sort_data[mode]=relevancy_monthly_grouped&media_type=all) and look for ads from the top spender. What do you notice about those ads? There's no code required here.

**A2.** Ashwani Jain (dem) ran for governor of MD, Peter Franchot  (dem) ran for election for Governor of Maryland. He lost in the Democratic primary on July 19, 2022. Monique Anderson-Walker, Franchot partner, elected to Lt. Governor in Prince George’s County, MD Council, District 8. Billy Boniface (rep) runs for Harford County Executive. Jon Baron (dem) ran for election for Governor of Maryland. He lost in the Democratic primary on July 19, 2022. The Working Families Party is building their own party on top of the two-party system in the United States — and its working. They organize outside the two parties, and then they recruit and train people-powered candidates up and down the ballot and run them to win. Former U.S. Education Secretary and Maryland gubernatorial candidate John King Jr. (dem) has tapped attorney and women’s rights advocate Michelle Siri as his choice for lieutenant governor. Eve T. Shuman (dem) ran for election to the Prince George's County Council, District 3 in Maryland. Shuman was on the ballot in the Democratic primary on July 19, 2022. John McCarthy, Democrat for Montgomery County State's Attorney. 

Its interesting how most candidates paid huge amounts on one ad to impress audiences. For instance, Ashwani Jain posted ad almost every day spending 200-300 for each ad. Whereas a person like Peter Franchot spent 15-20 k boxes on one ad. 


**Q3.** Given the list of top spenders in A1 and the primary results for governor (https://elections.maryland.gov/elections/2022/primary_results/gen_results_2022_1.html), who is missing from that Q1 list? Did those campaigns run Facebook ads? Who are the biggest recipients of media expenses by the Democratic primary winner for governor (HINT: that person has two committees you should account for)?

**A3.** Billy Boniface (yes), WFP National Pac - Non-Contribution Account (no), Shuman  Eve T. (yes), McCarthy  John J. (yes very few). 

Moore Miller For Maryland Slate (3549004.69). Their other committee name is Moore  Wes For Maryland

```{r}
dem_governor <- Expense %>%
  group_by(`Payee Name` , `Committee Name`) %>% 
summarise(total_amount=sum(`Amount($)`)) %>% 
  arrange(desc(total_amount))
  
```

**Q4.** The media expense data you cleaned also has a standardized purpose column. Write code to generate a list of the committees with the total amount of spending on online advertising. How does this list compare to the one from Q1 about top Facebook spenders? Of the top 5 committees who spent the most on Facebook ads from Q1, are there any that appear to have spent *all* of their online advertising dollars on Facebook?

**A4.** They are absolutely different. Here the top five include:
Maryland REALTORS Political Fund (567234.18), Baron-Williams Slate for Maryland (547500.07), NATIONAL ASSOCIATION OF REALTORS FUND
(462285.62), Perez  Tom for Maryland (427539.05), Republican Governors Association (330907.88).

Those who won the election might spent that amount, but I am not sure about those who lost. For instance, Billy Boniface, has so many FB ads which are from last year too. But I don't think those ads were for this year's election, to be honest, it's confusing...


```{r}
online_ad <- payee_fb %>% 
 filter(`Expense Purpose` == "Online Advertising") %>% 
group_by(`Committee Name`) %>% 
summarise(total_amount=sum(`Amount($)`)) %>% 
  arrange(desc(total_amount))
```

**Q5** You've been asked to summarize your findings about spending on Facebook by Maryland political candidates. How much do we know and what don't we know? What's the best story idea you've found here?

**A5** One story idea that comes to my mind is that the amount spent by the top five committee is huge. For instance, Jain  Ashwani spent this amount, 62357.38, on Facebook ads, but when I check the FB ad, the amount spent on each ad is not that much. It costs like $200-300. Now, since he posted lots of ads, the amount might be right for him, for other candidates, I am quite unsure. 
