---
title: "lab_01"
author: "Derek Willis"
date: "7/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
wv_winred_contribs <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** 13,382 rows, 24 columns

```{r}
glimpse(wv_winred_contribs)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
**A2** Martinsburg, 781

```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    )%>%
  arrange(desc(count_contribs))
  
```
```
**Q3** Which West Virginia city accounted for the fewest contributions? How many?
**A3** It kept giving errors. I tried both "asc" and "ascending", didn't work. A reason for that might be there are many cities in West Virginia with one contribution, and it's not just one city, so...

```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    )%>%
  arrange(asc(count_contribs))
```
```{r}
```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    )%>%
  arrange(asc(count_contribs))
```


```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    )%>%
  arrange(ascending(count_contribs))
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** Martinsburg is a state with major republicans elected and won elections. A reason I guess is that republicans invested and didn't want democrats to win.

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?
**A5** Earmarked for SAVE AMERICA JOINT FUNDRAISING COMMITTEE (C00770941). 517, and the average is 25

```{r}
wv_winred_contribs %>%
  group_by(memo_text, amount) %>%
  summarise(
    count_contribs = n(),
    mean_amount = mean(amount)
  ) %>% 
 arrange(desc(count_contribs))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** SAVE AMERICA JOINT FUNDRAISING COMMITTEE. If I am not mistaken, the name is taken from Save American political action committee established by Donald Trump.

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by city):
* count of contributions
* total amount of all contributions
* the average contribution amount
* the median contribution amount
* the smallest contribution amount
* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**. I tried to save it as a variable object called 'wv_summary' but it didn't work, I don't know what I did wrong but the wv_summary was saved as as RData in the window below environment. One of the story ideas came to my mind is why so much investment happened by Republicans in Martinsburg? Was it because they wanted to make sure a republican would win the election? Maybe, but interesting.
Martinsburg had the largest contribution but Charleston had larger amount gathered, interesting too. The min amount collected in Martinsburg is 0.05 and the max amount collected is 500. The median is 10,000. 

```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))

```
